---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%", dpi = 100, fig.retina = 2
)
```

# ProbBreed

<!-- badges: start -->
[![Lifecycle: stable](https://img.shields.io/badge/lifecycle-stable-brightgreen.svg)](https://lifecycle.r-lib.org/articles/stages.html#stable)
[![CRAN status](https://www.r-pkg.org/badges/version/ProbBreed)](https://CRAN.R-project.org/package=ProbBreed)
[![ProbBreed status badge](https://saulo-chaves.r-universe.dev/badges/ProbBreed)](https://saulo-chaves.r-universe.dev/ProbBreed)
[![CRAN downloads](https://cranlogs.r-pkg.org/badges/ProbBreed)](https://cran.r-project.org/package=ProbBreed)
<!-- badges: end -->

ProbBreed employs Bayesian statistics to analyse multi-environment trials' data, and uses its outputs to estimate the marginal and pairwise probabilities of superior performance and superior stability of the genotypes, as well as their conditional probability of superior performance. The method is thoroughly described at https://doi.org/10.1007/s00122-022-04041-y.
 
## Installation

You can install the CRAN version of `ProbBreed` using the following command:
```r
install.packages("ProbBreed")
```

Alternatively, you can install the development version of `ProbBreed` from [GitHub](https://github.com/saulo-chaves/ProbBreed) with:

``` r
# install.packages("devtools")
devtools::install_github("saulo-chaves/ProbBreed")
```

## Usage

```{r echo=TRUE}
library(ProbBreed)
```

Currently, `ProbBreed` has nine available models implemented in the `bayes_met` function. See `?bayes_met` for more details. An examples using the `soy` example dataset is described below:

```{r eval=TRUE, echo=TRUE, results='hide'}
model = bayes_met(data = soy,
                gen = "Gen",
                loc = "Loc",
                repl = NULL,
                year = NULL,
                reg = NULL,
                res.het = TRUE,
                trait = 'Y',
                iter = 6000, cores = 4, chains = 4)
```

`gen`, `loc`, `repl`, `year` and `reg` are all column names that contain information on genotypes, environments (locations), replicates, years (or seasons) and regions (or mega-environments). The `soy` dataset has only adjusted means, so only `gen` and `loc` have values. `res.het` indicates if a per-environmental residual variance should be estimated. `trait` is the column in `data` that contain the phenotypic observations. The other arguments are specifications for model fitting: the number of iterations, cores and chains. Feel free to customize these and other options according to your necessity.

The output of this function will be an object of class `stanfit`, which should be used in the `extr_outs` function for further processing before computing the probabilities per se. This function also provides some useful diagnostics. Here is how to use it:

```{r eval=TRUE, echo=TRUE, results='hide', message=FALSE, warning=FALSE}
outs = extr_outs(model = model,
                 probs = c(0.05, 0.95),
                 verbose = TRUE)
```

The object of class `extr` provided by this function contains the effects' posterior and maximum posterior, the models' variance components and some posterior predictive checks. Here are them:

```{r}
outs$variances
outs$ppcheck
```

You can also the `plot` S3 method for some useful visualizations. For e.g., the comparison between the empirical and sampled phenotype illustrates the model's convergence:

```{r}
plot(outs)
```

See `?plot.extr` for more details and further options.

After these two steps, everything is set to compute the probabilities. This can be done using the function `prob_sup`:


A basic workflow using the available data is:

```{r, eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
results = prob_sup(extr = outs, 
                   int = .2,
                   increase = TRUE, 
                   save.df = FALSE, 
                   verbose = TRUE)
```

This function generates an object of class `probsup`, which contains two lists: `across` and `within`. As their names suggest, the `across` list has the across-environments probabilities, and is suitable for a broader recommendation. Conversely, the `within` results are more appropriate to specific recommendations. For example, here are some probability of superior performances across and within environments:

```{r}
head(results$across$perfo)
head(results$within$perfo$gl)
```

The S3 method `plot` is also available for `probsup` objects. Here are some of them:

-   Probability of superior performance across environments

```{r echo=TRUE}
plot(results)
```

-   Pairwise probability of superior performance across environments

```{r echo=TRUE}
plot(results, category = "pair_perfo", level = "across")
```

-   Probability of superior stability

```{r echo=TRUE}
plot(results, category = "stabi")
```

-   Probability of superior performance within environments

```{r echo=TRUE}
plot(results, category = "perfo", level = "within")
```

The grey cells are environments where a given genotype was not assessed. See more options at `?plot.probsup`. 

The estimation of these probabilities are strictly related to some key questions that constantly arises in plant breeding, like:

-   **What is the risk of recommending a selection candidate for a target population of environments?**

-   **What is the probability of a given selection candidate having good performance if recommended to a target population of environments? And for a specific environment?**

-   **What is the probability of a given selection candidate having better performance than a cultivar check in the target population of environments? And in specific environments?**

-   **How probable is it that a given selection candidate performs similarly across environments?**

-   **What are the chances that a given selection candidate is more stable than a cultivar check in the target population of environments?**

-   **What is the probability that a given selection candidate having a superior and invariable performance across environments?**

For a more detailed tutorial, see https://saulo-chaves.github.io/ProbBreed_site/.

## Citation

Thank you for using `ProbBreed`! Please, do not forget to cite: 

```{r eval=FALSE, echo=TRUE}
citation('ProbBreed')
```


