---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# ProbBreed

<!-- badges: start -->
<!-- badges: end -->

ProbBreed employs Bayesian statistics to analyze multi-environment trials' data, and uses its outputs to computate the marginal, pairwise, and conditional probability of superior performance of the genotypes. The method is thoroughly described at https://doi.org/10.1007/s00122-022-04041-y. This is a work in progress.

## Installation

You can install the development version of ProbBreed from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("saulo-chaves/ProbBreed")
```

## Usage

A basic workflow using the available data is:

```{r example, eval=FALSE, echo=TRUE}
library(ProbBreed)

mod = bayes_met(data = maize, 
        gen = c("Hybrid", "normal", "cauchy"), 
        env = c("Location", "normal", "cauchy"),
        rept = list(c("Rep", "normal", "cauchy"), c("Block", "normal", "cauchy")),
        reg = list(c("Region", "normal", "cauchy"), c("normal", "cauchy")),
        res.het = T,
        sigma.dist = c("cauchy", "cauchy"), mu.dist = c("normal", "cauchy"),
        gei.dist = c("normal", "normal"), trait = "GY", hyperparam = "default",
        iter = 100, cores = 2, chain = 2) 
# You may want to increase the number of iterations, cores and chains

outs = extr_outs(data = maize, trait = "GY", gen = "Hybrid", model = mod,
                 effects = c("r", "b", "l", "m", "g", "gl", "gm"),
                 nenv = 16, res.het = TRUE, check.stan.diag = TRUE)

margs = marg_prob(data = maize, trait = "GY", gen = "Hybrid", 
                       env = "Location", increase = T,
                       extr_outs = outs, int = .2, save.df = F, interactive = F)

conds = cond_prob(data = maize, trait = "GY", gen = "Hybrid",env = "Location", 
                  extr_outs = outs, reg = "Region", int = .2,increase = T,
                  save.df = F, interactive = F)


```


