---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# ProbBreed

<!-- badges: start -->
[![Lifecycle: stable](https://img.shields.io/badge/lifecycle-stable-brightgreen.svg)](https://lifecycle.r-lib.org/articles/stages.html#stable)
[![CRAN status](https://www.r-pkg.org/badges/version/ProbBreed)](https://CRAN.R-project.org/package=ProbBreed)
[![ProbBreed status badge](https://saulo-chaves.r-universe.dev/badges/ProbBreed)](https://saulo-chaves.r-universe.dev/ProbBreed)
<!-- badges: end -->

ProbBreed employs Bayesian statistics to analyse multi-environment trials' data, and uses its outputs to estimate the marginal and pairwise probabilities of superior performance and superior stability of the genotypes, as well as their conditional probability of superior performance. The method is thoroughly described at https://doi.org/10.1007/s00122-022-04041-y.
 
## Installation

You can install the CRAN version of `ProbBreed` using the following command:
```r
install.packages("ProbBreed")
```

Alternatively, you can install the development version of `ProbBreed` from [GitHub](https://github.com/saulo-chaves/ProbBreed) with:

``` r
# install.packages("devtools")
devtools::install_github("saulo-chaves/ProbBreed")
```

## Usage

A basic workflow using the available data is:

```{r, eval=FALSE, echo=TRUE}
library(ProbBreed)

mod = bayes_met(data = soy,
                gen = "Gen",
                env = "Env",
                repl = NULL,
                year = NULL,
                reg = "Reg",
                res.het = F,
                trait = "Y",
                iter = 2000, cores = 2, chains = 4)

outs = extr_outs(data = soy, trait = "Y", model = mod,
                 probs = c(0.05, 0.95),
                 check.stan.diag = FALSE, 
                 verbose = TRUE)

results = prob_sup(data = soy, trait = "Y", gen = "Gen", env = "Env",
                   mod.output = outs, reg = 'Reg', year = NULL, int = .2,
                   increase = TRUE, save.df = FALSE, interactive = FALSE, 
                   verbose = TRUE)

```


